### 2022-03-15 omicron-generation.tex
tweaking abstract

Estimating the differences in the incubation-period and serial-interval distributions of SARS-CoV-2 variants is critical to understanding changes in their transmission dynamics and therefore controlling their outbreaks.
However, these comparisons are inherently difficult due to their dependence on epidemic growth rates---for example, when an epidemic is growing exponentially, we are more likely to observe individuals who have been infected recently, decreasing the mean observed incubation period and thereby increasing the mean observed serial interval.
Here, we analyze serial-interval data describing transmissions of the Delta and Omicron variants from Netherlands at the end of December 2021, during which the number of infections caused by the Delta variant decreased as the number of Omicron infections increased. 
Accounting for growth-rate differences of two variants, we estimate similar mean incubation periods (3.8--4.5 days) for both variants but a shorter mean generation interval for the Omicron variant (3.3 days; 95\% CI: 2.8--3.7 days) than for the Delta variant (4.2 days; 95\% CI: 4.0--4.4 days).
These differences may reflect higher reproduction numbers of the Omicron variant, which can drive faster susceptible depletion among close contacts.
Neglecting differences in the generation-interval distributions can further bias in estimates of transmission advantages of the Omicron variant.

>>>>

Estimating the differences in the incubation-period and serial-interval distributions of SARS-CoV-2 variants is critical to understanding changes in their transmission dynamics and therefore controlling their outbreaks.
However, these comparisons are inherently difficult due to their dependence on epidemic growth rates---for example, when an epidemic is growing exponentially, a cohort of infected individuals who developed symptoms at the same time are more likely to have shorter incubation periods.
Here, we re-analyze incubation-period and serial-interval data collected by Backer \textit{et al.}, describing transmissions of the Delta and Omicron variants from Netherlands at the end of December 2021.
Previous analysis of the same data set reported shorter mean observed incubation period (3.2 days vs 4.4 days) and serial interval (3.5 days vs 4.1 days) for the Omicron variant, but the number of infections caused by the Delta variant decreased during this period as the number of Omicron infections increased.
When we account for growth-rate differences of two variants, we estimate similar mean incubation periods (3.8--4.5 days) for both variants but a shorter mean generation interval for the Omicron variant (3.3 days; 95\% CI: 2.8--3.7 days) than for the Delta variant (4.2 days; 95\% CI: 4.0--4.4 days).
The differences in generation intervals may reflect higher reproduction numbers of the Omicron variant, which can drive faster susceptible depletion among close contacts.
Neglecting differences in the generation-interval distributions can further bias in estimates of transmission advantages of the Omicron variant.
